### 2014-11-13 nqueens.tex
Several corrections. Reduce size of texttt.

Since computation goes from the top row to the bottom row, not all
placements of nodes to threads perform equally. This is especially
true because the bottom rows tend to perform the most work. The best
placement is then to split the board vertically so each thread gets
the same number of columns with axiom \texttt{set-cpu(A, vertical(X, Y)}, where
\texttt{X} and \texttt{Y} are the coordinates of the board. Our experiments show that, on a shared
memory system, it does not matter much if we start with a bad
placement since node stealing overcomes those issues by load balancing
dynamically.  However, the N Queens program incrementally builds and shares lists representing
valid board states that are transmitted from top to bottom.
In order to improve memory locality we should then manipulate board
states that share a significant number of elements because each board state
needs to be iterated before extended with a new position.
To accomplish this, we used coordination axioms \texttt{set-default-priority(A,
X)}, where \texttt{X} is the row number (0 is the top row) in order to
prioritize the evaluation of squares closer to the bottom of the board.

Experimental results are presented in Fig.~\ref{results:nqueens}.  We
first ran the coordinated program using dynamic partitioning (with
work stealing) and saw some performance improvements
(\textbf{Coordinated}). However, using dynamic partitioning is bad for
memory locality since threads will often steal nodes.  We then
added \texttt{set-static(A)} as a coordination axiom and saw
performance improvement (\textbf{Static}), especially when the board
columns are perfectly partitioned by the number of threads (in our
case 13 threads). We also analyzed N Queens execution with Valgrind's
CacheGrind~\cite{Nethercote:2007:VFH:1250734.1250746} and found that the
\texttt{Static} version has half the number of data caches misses (last level)
of the \texttt{Regular} version and a slight more than half of the
\texttt{Coordinated} version. This validates our hypothesis about the
improvement seen in the \texttt{Static} version.

\begin{topfig}
   \begin{center}
      \includegraphics[width=5.5cm]{results/8queens-13.png}
   \end{center}
   \scap{results:nqueens}{Experimental results for the 13 Queens program.
   Coordination works best when 1 column is assigned to one thread and bottom
   rows have higher priority.}
\end{topfig}

>>>>

Since computation goes from the top row to the bottom row, not all
placements of nodes to threads perform equally. This is especially
true because the bottom rows tend to perform the most work. The best
placement is then to split the board vertically so each thread gets
the same number of columns with axiom \mytt{set-cpu(A, vertical(X, Y)}, where
\mytt{X} and \mytt{Y} are the coordinates of the board. Our experiments show that, on a shared
memory system, it does not matter much if we start with a bad
placement since node stealing overcomes those issues by load balancing
dynamically.  However, the N Queens program incrementally builds and shares lists representing
valid board states that are transmitted from top to bottom.
In order to improve memory locality we should then manipulate board
states that share a significant number of elements because each board state
needs to be iterated before extended with a new position.
To accomplish this, we used coordination axioms \mytt{set-default-priority(A,
X)}, where \mytt{X} is the row number (0 is the top row) in order to
prioritize the evaluation of squares closer to the bottom of the board.

Experimental results are presented in Fig.~\ref{results:nqueens}.  We
first ran the coordinated program using dynamic partitioning (with
work stealing) and saw some performance improvements
(\textbf{Coordinated}). However, using dynamic partitioning is bad for
memory locality since threads will often steal nodes.  We then
added \mytt{set-static(A)} as a coordination axiom (line
\textbf{Static/Regular}) and saw a significant drop in data cache misses and an
improvement in performance, specially when using 13 threads (the number of
columns).

\begin{topfig}
   \begin{center}
      \includegraphics[width=\figsize]{results/8queens-13.png}
   \end{center}
   \scap{results:nqueens}{Experimental results for the 13 Queens program.
   Coordination works best when 1 column is assigned to one thread and bottom
   rows have higher priority.}
\end{topfig}

### 2014-11-13 related.tex
Several corrections. Reduce size of texttt.

Linda~\cite{linda} is arguably the most famous coordination model. Linda
implements a data-driven coordination model and features a \emph{tuple space}
that can be manipulated using the following coordination directives:
\texttt{out(t)} to write a tuple \texttt{t} into the tuple space; \texttt{in(t)}
to read a tuple using the template \texttt{t}; \texttt{rd(t)} to retrieve a copy
of the tuple \texttt{t} from the tuple space; and \texttt{eval(p)} to add a
process \texttt{p} in the tuple space and execute it in parallel.  Linda was
implemented on top of many popular languages by simply creating a communication
and storage mechanism for the tuple space and then adding the directives as a
language library.

>>>>

Linda~\cite{linda} is arguably the most famous coordination model. Linda
implements a data-driven coordination model and features a \emph{tuple space}
that can be manipulated using the following coordination directives:
\mytt{out(t)} to write a tuple \mytt{t} into the tuple space; \mytt{in(t)}
to read a tuple using the template \mytt{t}; \mytt{rd(t)} to retrieve a copy
of the tuple \mytt{t} from the tuple space; and \mytt{eval(p)} to add a
process \mytt{p} in the tuple space and execute it in parallel.  Linda was
implemented on top of many popular languages by simply creating a communication
and storage mechanism for the tuple space and then adding the directives as a
language library.

### 2014-11-13 splash-bp.tex
Several corrections. Reduce size of texttt.

\begin{tightdescription}
   \item[Tree building]: Each node has a \texttt{inactive} fact that is used to
   start the tree building process. When the highest priority node is picked, a
   \texttt{tree} is created that will navigate through the tree. In lines 15-21,
   we use an \emph{aggregate}~\cite{cruz-iclp14} to gather all the neighbor
   nodes that have a positive priority (due to a new belief update) and are in the
   same thread. Nodes are collected into list \texttt{L} and
   appended to list \texttt{Next} (line 21).

   \item[First phase]: When the number of nodes in the tree reaches a certain
   limit, a \texttt{first-phase} (lines 11-12) is generated to update the beliefs of
   all nodes in the tree. As the nodes are updated, starting from the leaves and
   ending at the root, an \texttt{update} fact is derived to update the belief
   values (line 35).

@@

The \texttt{set-static} and \texttt{set-cpu} action facts are used in
line 2 to (1) force nodes to stay in the thread and (2) 
partition nodes as a grid of threads. This sets up well defined areas
of nodes for threads to build splash trees on.

@@

// SECOND PHASE
second-phase(A, [], _)
   -o *underline{remove-priority(A)}, inactive(A), update(A).
second-phase(A, [A], Back)
   -o update(A), inactive(Back),
      inactive(A), *underline{remove-priority(A)}.
second-phase(A, [A, B | Next], Back)
   -o update(A), inactive(Back), *underline{schedule-next(B)},
      second-phase(B, [B | Next], A).
\end{Verbatim}
  \scap{code:sbp}{Coordination code for the Splash Belief Propagation program. 
%CLM needs 50 lines of rules to implement splash trees, while GraphLab needs a total of 350 lines of C++ to implement the same functionality.
  Note
     that when a linear fact is prefixed by \texttt{@}, it indicates that the
     fact is going to be re-derived.}
\end{topfig}
\normalsize

\begin{dblfig}
   \begin{center}
      \subfloat[]{\includegraphics[width=5.5cm]{results/system_belief-propagation-400.png}}
      \subfloat[]{\includegraphics[width=5.5cm]{results/system_splash-bp-400.png}}
      \subfloat[]{\includegraphics[width=5.5cm]{results/system_improve_belief-propagation-400.png}}
   \end{center}
   \scap{results:splash_bp}{Experimental results for LBP and SBP. (a) shows the
      scalability of LBP for both CLM and GraphLab and (b) shows the
      scalability of SBP. (c) presents the
      improvements seen in SBP against LBP, where SBP runs, on average, 1.5 to 2.5
      times faster than LBP.}
\end{dblfig}

>>>>

\begin{tightdescription}
   \item[Tree building]: Each node has a \mytt{inactive} fact that is used to
   start the tree building process. When the highest priority node is picked, a
   \mytt{tree} is created that will navigate through the tree. In lines 15-21,
   we use an \emph{aggregate}~\cite{cruz-iclp14} to gather all the neighbor
   nodes that have a positive priority (due to a new belief update) and are in the
   same thread. Nodes are collected into list \mytt{L} and
   appended to list \mytt{Next} (line 21).

   \item[First phase]: When the number of nodes in the tree reaches a certain
   limit, a \mytt{first-phase} (lines 11-12) is generated to update the beliefs of
   all nodes in the tree. As the nodes are updated, starting from the leaves and
   ending at the root, an \mytt{update} fact is derived to update the belief
   values (line 35).

@@

The \mytt{set-static} and \mytt{set-cpu} action facts are used in
line 2 to (1) force nodes to stay in the thread and (2) 
partition nodes as a grid of threads. This sets up well defined areas
of nodes for threads to build splash trees on.

@@

// SECOND PHASE
second-phase(A, [], _)
   -o *underline{remove-priority(A)}, inactive(A), update(A).
second-phase(A, [A], Back)
   -o update(A), inactive(Back),
      inactive(A), *underline{remove-priority(A)}.
second-phase(A, [A, B | Next], Back)
   -o update(A), inactive(Back), *underline{schedule-next(B)},
      second-phase(B, [B | Next], A).
\end{Verbatim}
  \scap{code:sbp}{Coordination code for the Splash Belief Propagation program. 
%CLM needs 50 lines of rules to implement splash trees, while GraphLab needs a total of 350 lines of C++ to implement the same functionality.
  Note
     that when a linear fact is prefixed by \mytt{@}, it indicates that the
     fact is going to be re-derived.}
\end{topfig}
\normalsize

\begin{dblfig}
   \begin{center}
      \subfloat[]{\includegraphics[width=\figsize]{results/system_belief-propagation-400.png}}
      \subfloat[]{\includegraphics[width=\figsize]{results/system_splash-bp-400.png}}
      \subfloat[]{\includegraphics[width=\figsize]{results/system_improve_belief-propagation-400.png}}
   \end{center}
   \scap{results:splash_bp}{Experimental results for LBP and SBP. (a) shows the
      scalability of LBP for both CLM and GraphLab and (b) shows the
      scalability of SBP. (c) presents the
      improvements seen in SBP against LBP, where SBP runs, on average, 1.5 to 2.5
      times faster than LBP.}
\end{dblfig}

### 2014-11-13 sssp.tex
Several corrections. Reduce size of texttt.

We start by adding coordination to the SSSP program described before.
The coordinated version of the
SSSP~(Fig.~\ref{code:shortest_path_program_coord}) uses the coordination fact
\texttt{set-priority} (line 14) and a global program directive to order
priorities in ascending order (line 5).

@@

The most interesting property of the SSSP program presented in
Fig.~\ref{code:shortest_path_program_coord} is that it remains provably correct,
although it applies rules using smarter ordering and the code remains
declarative. The derivation of \texttt{set-priority} does not change the
behavior of the logical rules.

@@

\begin{topfig}
\vspace*{-5ex}
   \begin{center}
      \includegraphics[width=5cm]{results/shortest-uspowergrid.png}
   \end{center}
   \scap{results:sssp_uspowergrid}{Scalability results for SSSP run on the 
   US power grid network. The speedup values are computed using the
   execution time for 1 thread.}
\vspace*{-3ex plus 0pt minus 1ex}
\end{topfig}

>>>>

We start by adding coordination to the SSSP program described before.
The coordinated version of the
SSSP~(Fig.~\ref{code:shortest_path_program_coord}) uses the coordination fact
\mytt{set-priority} (line 14) and a global program directive to order
priorities in ascending order (line 5).

@@

The most interesting property of the SSSP program presented in
Fig.~\ref{code:shortest_path_program_coord} is that it remains provably correct,
although it applies rules using smarter ordering and the code remains
declarative. The proof of correctness considers that, eventually, the shortest
path is computed at all nodes of the graph until all shortest paths are found.
The derivation of \mytt{set-priority} does not change this
proof.

@@

\begin{topfig}
\vspace*{-5ex}
   \begin{center}
      \includegraphics[width=\figsize]{results/shortest-uspowergrid.png}
   \end{center}
   \scap{results:sssp_uspowergrid}{Scalability results for SSSP run on the 
   US power grid network. The speedup values are computed using the
   execution time for 1 thread.}
\vspace*{-3ex plus 0pt minus 1ex}
\end{topfig}
